---
title: "2026 *Ulva australis* Dessiccation analysis"
title-block-banner: true
date: today
date-format: "YYYY MMM MM"
author: 
  - name: Dai Sumiyoshi
  - name: Kaiki Tajiri
  - name: Masahiro Suzuki
  - name: Gregory N. Nishihara
  - name: Ryuta Terada
execute: 
  echo: false
  message: false
  warning: false
---

```{r}
#| include: false
library(tidyverse)
library(ggpubr)
library(ggtext)
library(magick)
library(patchwork)
library(showtext)
library(brms)
library(tidybayes)
library(gt)
library(gnnlab)
library(utils)
library(readxl)
library(future)
library(bayestestR)

```

```{r}
#| include: false
font_add_google("Noto Sans", family = "ns")
theme_pubr(10, "ns", border = TRUE) |> 
  set_theme()
theme_pubr() |> 
  replace_theme(
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    axis.title.x = element_markdown(family = "ns", margin = margin(0.5, 0, 0.5, 0, "lines")),
    axis.title.y = element_markdown(family = "ns", angle= 90, margin = margin(0, 0.5, 0, 0.5, "lines")),
  )
showtext_auto()
folder = "~/Dropbox/MS_Terada_Greg/"
folder = "./Data/"
fnames = dir(folder, recursive = TRUE)
```


```{r}
ftext = "1st"
fname = str_c(folder, str_subset(fnames, ftext)) 
data = read_xlsx(fname, range = "A1:P170") |> drop_na()
data = data |> select(state = Status, minutes = Min, N = `#`, period = `...4`, matches("^Y"))
data = data |> mutate(period = str_extract(period, "[0-9]+(?=m)")) |> 
  pivot_longer(matches("^Y")) |> 
  mutate(ID = as.numeric(str_extract(name, "[0-9]+")))
data2 = read_xlsx(fname, range = "A1:D180") |> drop_na()
data2 = data2 |> filter(str_detect(Status, "Desicc"))
data3 = read_xlsx(fname, range = "V1:AA180") |> drop_na()
data23 = 
  bind_cols(data2, data3) |> 
   select(minutes = Min, N = `#`, period = `...4`, 
          fw0 = matches("Initial"),
          fw1 = matches("Fresh W"),
          foil = matches("Foil"),
          total = matches("Sample")) |> 
  mutate(dw = total - foil) |> 
  mutate(awc = (fw1 - dw) / (fw0 - dw) * 100) |> 
  mutate(period = str_extract(period, "[0-9]+(?=m)"))
data = full_join(data, data23, by = join_by(minutes, N, period))

odata = data
data = odata |> filter(!str_detect(state, "Init")) |> 
  mutate(state = factor(state, levels = c("Desiccation", "30-min rehydration", "1-d rehydration")))
```

```{r}
xlabel = "Time of exposure (min)"
ylabel = "Effective quantum yield (&Delta;F/Fm&prime;)"
xbreaks = unique(data$minutes)
xbreaks = seq(min(xbreaks), max(xbreaks), by = 20)
# 1. Create a label dataframe 
# We take distinct categories and generate the "A: ..." label here
labeldf = data |> expand(state, x = min(minutes), y = 1) |> 
  mutate(
    l = str_glue("({LETTERS[seq_along(state)]}) {state}")
  )

p1 = ggplot(data) + 
  geom_point(aes(x = minutes, y = value), 
             position = position_jitter(width = 0.2), 
             shape = 21,
             fill = "black",
             stroke = 0,
             alpha = 0.8) +
  geom_richtext(aes(x = x, y = y, label = l),
                data = labeldf,
    vjust = 1, hjust = 0,
    color = "transparent",
    fill = "transparent",
    text.color = "black",
    family = "ns",
    size = 3
  ) +
  scale_x_continuous(xlabel, breaks = xbreaks, limits = range(xbreaks)) +
  scale_y_continuous(ylabel, breaks = seq(0, 1, by = 0.25), limits = c(0, 1)) +
  facet_wrap(vars(state), ncol = 1) +
  theme(
    strip.text = element_blank(),
    strip.background = element_blank()
  );p1

pdfname = "PAM-Desiccation-minutes-plot.pdf"
save_plot(pdfname, plot = p1, width = 80, height = 80*1.5)
```



::: {#fig-pam-desiccation-01}


```{r}
image_read(str_replace(pdfname, "pdf", "png"))
```

The chronological response of the effective quantum yield (&Delta;F/Fm&prime;) in the 
green alga, *Ulva australis* from Nagashima Island, Kagoshima, Japan to (A) only air 
exposure, (B) air exposure and subsequent 30-minute rehydration, and (C) 
air exposure and subsequent 1-day rehydration conducted 
at 20&deg;C and 50% humidity under the dim light (20 &mu;mol 
photons m^-2^ s^-1^). 
The points are jittered horizontally for clarity (n = 50).

:::


```{r}
xlabel = "Relative water content (%)"
ylabel = "Effective quantum yield (&Delta;F/Fm&prime;)"
xbreaks = unique(data$awc)
xbreaks = seq(0, 100, by = 20)
# 1. Create a label dataframe 
# We take distinct categories and generate the "A: ..." label here
labeldf = data |> expand(state, x = 0, y = 1) |> 
  mutate(
    l = str_glue("({LETTERS[seq_along(state)]}) {state}")
  )

p1 = ggplot(data) + 
  geom_point(aes(x = awc, y = value), 
             position = position_jitter(width = 0.2), 
             shape = 21,
             fill = "black",
             stroke = 0,
             alpha = 0.8) +
  geom_richtext(aes(x = x, y = y, label = l),
                data = labeldf,
    vjust = 1, hjust = 0,
    color = "transparent",
    fill = "transparent",
    text.color = "black",
    family = "ns",
    size = 3
  ) +
  scale_x_continuous(xlabel, breaks = xbreaks, limits = range(xbreaks)) +
  scale_y_continuous(ylabel, breaks = seq(0, 1, by = 0.25), limits = c(0, 1)) +
  facet_wrap(vars(state), ncol = 1) +
  theme(
    strip.text = element_blank(),
    strip.background = element_blank()
  );p1

pdfname = "PAM-Desiccation-RWC-plot.pdf"
save_plot(pdfname, plot = p1, width = 80, height = 80*1.5)
```

::: {#fig-pam-dessication-02}


```{r}
image_read(str_replace(pdfname, "pdf", "png"))
```

The response of the effective quantum yield (&Delta;F/Fm&prime;) in the 
green alga, *Ulva australis* from Nagashima Island, Kagoshima, Japan along the relative water content (%RWC) 
to (A) only air exposure, 
(B) air exposure and subsequent 30-minute rehydration, and 
(C) air exposure and subsequent 1-day rehydration conducted 
at 20&deg;C and 50% humidity under the dim light (20 &mu;mol 
photons m^-2^ s^-1^). 
The points are jittered horizontally for clarity (n = 50).

:::


```{r}
cores = chains = 5
seed = 2026
warmup = 4000
iterations = warmup + 10000
file_refit = "never"
control = list(adapt_delta = 0.995, max_treedepth = 15)
filename_null = "pam-desiccation-null-model"
filename_full = "pam-desiccation-full-model"

nullmodel = bf(value ~ s(awc, k = 8, bs = "cr"), zi ~ s(awc, k = 8, bs = "cr")) + zero_inflated_beta()
fullmodel = bf(value ~ s(awc, k = 8, bs = "cr") + s(awc, k = 8, bs = "cr", by = state) + state, zi ~ s(awc, k = 8, bs = "cr") + s(awc, k = 8, bs = "cr", by = state) + state) + zero_inflated_beta()
# get_prior(nullmodel, data)
# get_prior(fullmodel, data)
priors_null = c(
  prior(student_t(3, 0, 2.5), class = Intercept),
  prior(student_t(3, 0, 2.5), class = b),
  prior(student_t(3, 0, 2.5), class = sds),
  prior(student_t(3, 0, 2.5), dpar = zi, class = Intercept),
  prior(student_t(3, 0, 2.5), dpar = zi, class = b),
  prior(student_t(3, 0, 2.5), dpar = zi, class = sds),
  prior(gamma(0.01, 0.01), class = phi)
)
priors_full = c(
  prior(student_t(3, 0, 2.5), class = Intercept),
  prior(student_t(3, 0, 2.5), class = b),
  prior(student_t(3, 0, 2.5), class = sds),
  prior(student_t(3, 0, 2.5), dpar = zi, class = Intercept),
  prior(student_t(3, 0, 2.5), dpar = zi, class = b),
  prior(student_t(3, 0, 2.5), dpar = zi, class = sds),
  prior(gamma(0.01, 0.01), class = phi)
)
```


```{r}
#| label: nullmodel

plan(cluster, workers = "mimir")
futurenull = future({
  brm(
    nullmodel,
    data = data,
    prior = priors_null,
    chains = chains,
    cores = cores,
    seed = seed,
    iter = iterations,
    warmup = warmup,
    refresh = 0,
    file = filename_null,
    file_refit = file_refit,
    backend = "cmdstanr",
    threads = threading(4),
    control = control,
    save_pars = save_pars(all = TRUE)
  )
})

z = 0
while(!resolved(futurenull)) {
  z0 = lubridate::now()
  message("Model is still sampling... ", z0)
  Sys.sleep(10) # Check every 10 seconds
  z = z + 10
}

message("Done! You can now run: fit <- value(futurenull)\n", "Total time is: ", z/60, " minutes.")
plan(sequential)
```

```{r}
bout_null = value(futurenull)

require(ssh)
remote = str_c("/home/gnishihara/MS_Terada_Nishihara_ulva_australis_202601/", filename_null, ".rds") |> print()
session = ssh_connect("mimir")
scp_download(session, files = remote, to = "./")
ssh_disconnect(session)
```


```{r}
#| label: fullmodel

plan(cluster, workers = "mimir")
futurefull = future({
  brm(
    fullmodel,
    data = data,
    prior = priors_full,
    chains = chains,
    cores = cores,
    seed = seed,
    iter = iterations,
    warmup = warmup,
    refresh = 0,
    file = filename_full,
    file_refit = file_refit,
    backend = "cmdstanr",
    threads = threading(4),
    control = control,
    save_pars = save_pars(all = TRUE)
  )
})

z = 0
while(!resolved(futurefull)) {
  z0 = lubridate::now()
  message("Model is still sampling... ", z0)
  Sys.sleep(10) # Check every 10 seconds
  z = z + 10
}

message("Done! You can now run: fit <- value(futurefull)\n", "Total time is: ", z/60, " minutes.")
plan(sequential)
```

```{r}
bout_full = value(futurefull)

require(ssh)
remote = str_c("/home/gnishihara/MS_Terada_Nishihara_ulva_australis_202601/", filename_full, ".rds") |> print()
session = ssh_connect("mimir")
scp_download(session, files = remote, to = "./")
ssh_disconnect(session)
```


