---
title: "2026 *Ulva australis* Dessiccation analysis"
title-block-banner: true
date: today
date-format: "YYYY MMM MM"
author: 
  - name: Dai Sumiyoshi
  - name: Kaiki Tajiri
  - name: Masahiro Suzuki
  - name: Gregory N. Nishihara
  - name: Ryuta Terada
execute: 
  echo: false
  message: false
  warning: false
  cache: true
---

```{r}
#| include: false
library(tidyverse)
library(ggpubr)
library(ggtext)
library(magick)
library(patchwork)
library(showtext)
library(brms)
library(tidybayes)
library(gt)
library(gnnlab)
library(utils)
library(readxl)
library(future)
library(bayestestR)

```

```{r}
#| include: false
font_add_google("Noto Sans", family = "ns")
theme_pubr(10, "ns", border = TRUE) |> 
  set_theme()
theme_pubr() |> 
  replace_theme(
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    axis.title.x = element_markdown(family = "ns", margin = margin(0.5, 0, 0.5, 0, "lines")),
    axis.title.y = element_markdown(family = "ns", angle= 90, margin = margin(0, 0.5, 0, 0.5, "lines")),
  )
showtext_auto()
folder = "~/Dropbox/MS_Terada_Greg/"
folder = "./Data/"
fnames = dir(folder, recursive = TRUE)

svscaler = function(x, n) {
  (x * (n - 1) + 0.5) / n
}
```


```{r}
ftext = "1st"
fname = str_c(folder, str_subset(fnames, ftext)) 
data = read_xlsx(fname, range = "A1:P170") |> drop_na()
data = data |> select(state = Status, minutes = Min, N = `#`, period = `...4`, matches("^Y"))
data = data |> mutate(period = str_extract(period, "[0-9]+(?=m)")) |> 
  pivot_longer(matches("^Y")) |> 
  mutate(ID = as.numeric(str_extract(name, "[0-9]+")))
data2 = read_xlsx(fname, range = "A1:D180") |> drop_na()
data2 = data2 |> filter(str_detect(Status, "Desicc"))
data3 = read_xlsx(fname, range = "V1:AA180") |> drop_na()
data23 = 
  bind_cols(data2, data3) |> 
  select(minutes = Min, N = `#`, period = `...4`, 
         fw0 = matches("Initial"),
         fw1 = matches("Fresh W"),
         foil = matches("Foil"),
         total = matches("Sample")) |> 
  mutate(dw = total - foil) |> 
  mutate(awc = (fw1 - dw) / (fw0 - dw) * 100) |> 
  mutate(period = str_extract(period, "[0-9]+(?=m)"))
data = full_join(data, data23, by = join_by(minutes, N, period))

odata = data
data = odata |> filter(!str_detect(state, "Init")) |> 
  mutate(state = factor(state, levels = c("Desiccation", "30-min rehydration", "1-d rehydration"))) |>
  mutate(nstate = as.numeric(state)) |> 
  rename(individual = N,
         location = ID) |> 
  select(state, nstate, individual, location, minutes, value, awc) |> 
  group_by(state) |> 
  mutate(n = length(awc)) |> 
  mutate(sawc = svscaler(awc/100, n),
         svalue = svscaler(value, n)) |> 
  ungroup() 

data2 = 
  data |> 
  group_by(state, nstate, individual, minutes) |> 
  reframe(
    value_sd = sd(value),
    value_m = mean(value),
    awc = first(awc),
    sawc = first(sawc)) 
```


```{r}
xlabel = "Time of exposure (min)"
ylabel = "Effective quantum yield (&Delta;F/Fm&prime;)"
xbreaks = unique(data$minutes)
xbreaks = seq(min(xbreaks), max(xbreaks), by = 20)
# 1. Create a label dataframe 
# We take distinct categories and generate the "A: ..." label here
labeldf = data |> expand(state, x = min(minutes), y = 1) |> 
  mutate(
    l = str_glue("({LETTERS[seq_along(state)]}) {state}")
  )

p1 = ggplot(data2) + 
  geom_point(aes(x = minutes, y = value_m), 
             shape = 21,
             fill = "black",
             stroke = 0,
             alpha = 0.8) +
  geom_richtext(aes(x = x, y = y, label = l),
                data = labeldf,
                vjust = 1, hjust = 0,
                color = "transparent",
                fill = "transparent",
                text.color = "black",
                family = "ns",
                size = 3
  ) +
  scale_x_continuous(xlabel, breaks = xbreaks, limits = range(xbreaks)) +
  scale_y_continuous(ylabel, breaks = seq(0, 1, by = 0.25), limits = c(0, 1)) +
  facet_wrap(vars(state), ncol = 1) +
  theme(
    strip.text = element_blank(),
    strip.background = element_blank()
  )

pdfname = "PAM-Desiccation-minutes-plot.pdf"
save_plot(pdfname, plot = p1, width = 80, height = 80*1.5)
```



::: {#fig-pam-desiccation-01}


```{r}
image_read(str_replace(pdfname, "pdf", "png"))
```

The chronological response of the effective quantum yield (&Delta;F/Fm&prime;) in the 
green alga, *Ulva australis* from Nagashima Island, Kagoshima, Japan to (A) only air 
exposure, (B) air exposure and subsequent 30-minute rehydration, and (C) 
air exposure and subsequent 1-day rehydration conducted 
at 20&deg;C and 50% humidity under the dim light (20 &mu;mol 
photons m^-2^ s^-1^). 
The points are jittered horizontally for clarity (n = 50).

:::


```{r}
xlabel = "Relative water content (%)"
ylabel = "Effective quantum yield (&Delta;F/Fm&prime;)"
xbreaks = unique(data$awc)
xbreaks = seq(0, 100, by = 20)
# 1. Create a label dataframe 
# We take distinct categories and generate the "A: ..." label here
labeldf = data |> expand(state, x = 0, y = 1) |> 
  mutate(
    l = str_glue("({LETTERS[seq_along(state)]}) {state}")
  )

p1 = ggplot(data2) + 
  geom_point(aes(x = awc, y = value_m),
             shape = 21,
             fill = "black",
             stroke = 0,
             alpha = 0.8) +
  geom_richtext(aes(x = x, y = y, label = l),
                data = labeldf,
                vjust = 1, hjust = 0,
                color = "transparent",
                fill = "transparent",
                text.color = "black",
                family = "ns",
                size = 3
  ) +
  scale_x_continuous(xlabel, breaks = xbreaks, limits = range(xbreaks)) +
  scale_y_continuous(ylabel, breaks = seq(0, 1, by = 0.25), limits = c(0, 1)) +
  facet_wrap(vars(state), ncol = 1) +
  theme(
    strip.text = element_blank(),
    strip.background = element_blank()
  )

pdfname = "PAM-Desiccation-RWC-plot.pdf"
save_plot(pdfname, plot = p1, width = 80, height = 80*1.5)
```

::: {#fig-pam-dessication-02}


```{r}
image_read(str_replace(pdfname, "pdf", "png"))
```

The response of the effective quantum yield (&Delta;F/Fm&prime;) in the 
green alga, *Ulva australis* from Nagashima Island, Kagoshima, Japan along the relative water content (%RWC) 
to (A) only air exposure, 
(B) air exposure and subsequent 30-minute rehydration, and 
(C) air exposure and subsequent 1-day rehydration conducted 
at 20&deg;C and 50% humidity under the dim light (20 &mu;mol 
photons m^-2^ s^-1^). 
The points are jittered horizontally for clarity (n = 50).

:::

Smithson, M., & Verkuilen, J. (2006). A better lemon squeezer? Maximum-likelihood regression with beta-distributed dependent variables. Psychological Methods, 11(1), 54–71. https://doi.org/10.1037/1082-989X.11.1.54

```{r}
data2 = 
  data |> 
  group_by(state, nstate, individual, minutes) |> 
  reframe(
    value_sd = sd(value),
    value_m = mean(value),
    awc = first(awc)) |> 
  select(state, nstate, minutes, awc, individual, value_m, value_sd) |>
  mutate(
    nstate = factor(nstate, labels = c("D", "Thirty", "Day"), ordered = TRUE),
    individual = factor(individual)) 

data2 = 
  data2 |> 
  mutate(outlier = ifelse(between(awc, 20, 30) & str_detect(individual, "2"), TRUE, FALSE)) 

data2 = data2 |> 
  group_nest(state, individual) |> 
  mutate(id = 1:n()) |> 
  unnest(data)

data2 = data2 |> 
  mutate(y = (value_m * (n() - 1) + 0.5)/n(),
         x = awc / 100)

data2 = data2 |> filter(!outlier)


data2 |> 
  ggplot() +
  geom_point(aes(x = x, y = y,color = individual)) + 
  facet_grid(vars(nstate), vars(individual))

```


```{r}
cores = chains = 5
seed = 2026
warmup = 4000
iterations = warmup + 10000
file_refit = "never"
control = list(adapt_delta = 0.995, max_treedepth = 20)
filename_full = "pam-desiccation-full-model"

file.remove(str_c(filename_full, ".rds"))

library(mgcv)
library(gratia)

gam01 = gam(
  y ~ s(x, bs = "ps") + s(x, by = nstate, bs = "ps", m = 1) + s(id, bs = "re"),
  data = data2, family = betar(link = "logit")
)

draw(gam01)
appraise(gam01)
```


```{r}
fullmodel = bf(
  y ~ nstate + s(x, bs = "ps") + s(x, by = nstate, bs = "ps", m = 1) + (1|individual)
  ) + beta()


priors_full = c(
  prior(normal(0, 1.5), class = Intercept),
  prior(normal(0, 1.5), class = b),
  prior(exponential(2), class = sd),
  prior(exponential(2), class = sds),
  # prior(normal(0, 1.5), dpar = zi, class = Intercept),
  # prior(normal(0, 1.5), dpar = zi, class = b),
  prior(gamma(0.01, 0.01), class = phi)
)
# 
# bout_full =
#   brm(
#     fullmodel,
#     data = data2,
#     prior = priors_full,
#     chains = 5,
#     cores = 5,
#     seed = seed,
#     iter = iterations,
#     warmup = warmup,
#     refresh = 20,
#     file = filename_full,
#     file_refit = file_refit,
#     backend = "cmdstanr",
#     control = list(adapt_delta = 0.95, max_treedepth = 20),
#     save_pars = save_pars(all = TRUE)
#   )
# 
# 
# p1 = pp_check(bout_full, ndraws = 50, type = "dens_overlay_grouped", group = "state")
# p2 = pp_check(bout_full, ndraws = 500, type = "stat_grouped", stat = "median", group = "state")
# p3 = pp_check(bout_full, ndraws = 500, type = "stat_grouped", stat = "mad", group= "state")
# p4 = pp_check(bout_full, ndraws = 500, type = "stat_grouped", stat = "mean", group= "state")
# p5 = pp_check(bout_full, ndraws = 500, type = "stat_grouped", stat = "sd", group= "state")
# 
# p1 / p2 / p3 /p4 /p5
```


```{r}
#| label: fullmodel


cores = chains = 5
seed = 2026
warmup = 4000
iterations = warmup + 10000
file_refit = "never"
control = list(adapt_delta = 0.995, max_treedepth = 20)
filename_full = "pam-desiccation-full-model"


file.remove(str_c(filename_full, ".rds"))
file.exists(str_c(filename_full, ".rds"))

if(!file.exists(str_c(filename_full, ".rds"))) {
  plan(cluster, workers = "mimir")
  futurefull = future({
    brm(
      fullmodel,
      data = data2,
      prior = priors_full,
      chains = chains,
      cores = cores,
      seed = seed,
      iter = iterations,
      warmup = warmup,
      refresh = 0,
      backend = "cmdstanr",
      control = control,
      save_pars = save_pars(all = TRUE)
    )
  }, seed = seed)
  
  z = 0
  sleeptime = 5
  while(!resolved(futurefull)) {
    z0 = lubridate::now()
    message("Model is still sampling... ", z0)
    Sys.sleep(sleeptime) # Check every 10 seconds
    z = z + sleeptime
  }
  
  message("Done! You can now run: fit <- value(futurefull)\n", "Total time is: ", z/60, " minutes.")
  bout_full = value(futurefull)
  bout_full |> write_rds(str_c(filename_full, ".rds"))
  plan(sequential); gc()
  # future({Sys.info()[["nodename"]]}) |> value()
} else {
  bout_full = read_rds(str_c(filename_full, ".rds"))
}
```


# Posterior predictive checks

```{r}
p1 = pp_check(bout_full, type = "dens_overlay_grouped",  group = "nstate", ndraws = 50)
p2 = pp_check(bout_full, type = "stat_grouped", stat = c("median"), ndraws = 500, binwidth = 0.01, group = "nstate")
p3 = pp_check(bout_full, type = "stat_grouped", stat = c("mean"),   ndraws = 500, binwidth = 0.01, group = "nstate")
p4 = pp_check(bout_full, type = "stat_grouped", stat = c("mad"),    ndraws = 500, binwidth = 0.01, group = "nstate")
p5 = pp_check(bout_full, type = "stat_grouped", stat = c("sd"),     ndraws = 500, binwidth = 0.01, group = "nstate")

pout = p1 / p2 / p3 / p4 / p5 

pdfname = "PAM-Desiccation-RWC-diag-plot.pdf"
save_plot(pdfname, plot = pout, width = 80*3, height = 80*2)
ce_df = conditional_effects(bout_full)
```

```{r}
xlabel = "Relative water content (%)"
ylabel = "Effective quantum yield (&Delta;F/Fm&prime;)"

labeldf = data2 |> expand(nesting(state, nstate), x = 0, y = 1) |> 
  mutate(
    l = str_glue("({LETTERS[seq_along(nstate)]}) {state}")
  )


p1 = ggplot() + 
  geom_richtext(aes(x = x, y = y, label = l),
                data = labeldf,
                vjust = 1, hjust = 0,
                color = "transparent",
                fill = "transparent",
                text.color = "black",
                family = "ns",
                size = 3
  ) +
  geom_ribbon(
    aes(
      x = x, ymin = lower__, ymax = upper__, 
    ),
    data = ce_df$`x:nstate`,
    alpha = 0.5
  ) +
  # geom_ribbon(
  #   aes(
  #     x = awc, ymin = .lower, ymax = .upper, fill = state
  #   ),
  #   data = df_out80,
  #   alpha = 0.5
  # ) +
  geom_line(
    aes(
      x = x, y = estimate__, 
    ),
    data = ce_df$`x:nstate`,
    alpha = 0.5
  ) +
  geom_point(aes(x = x, y = value_m,  shape = factor(individual)),
             data = data2) +
  scale_color_viridis_d("Treatment", end = 0.8,
                        labels = c("Desiccation", "30-min rehydration", "1-day rehydration"),
                        aesthetics = c("fill", "color")) +
  scale_shape_manual("Individual",values = c(0, 1, 2, 15, 16, 17)) +
  scale_x_continuous(xlabel, 
                     limits = c(0, 1),
                     breaks = seq(0, 1, by = .20),
                     labels = seq(0, 100, by = 20)) +
  scale_y_continuous(ylabel, limits = c(0,1),
                     breaks = seq(0, 1, by = 0.2)) +
  guides(shape = guide_legend(direction = "horizontal", ncol = 3, ),
         fill = "none",
         color = "none"
         ) +
  facet_wrap(vars(nstate), ncol = 1) +
  theme(
    legend.byrow = TRUE,
    legend.title = element_blank(),
    legend.text = element_markdown(size = rel(0.8)),
    legend.key.size = unit(5, "mm"),
    legend.position = "inside",
    legend.position.inside = c(1, 0),
    legend.justification = c(1, 0),
    legend.background = element_blank(),
    legend.margin = margin(0, 4, 0, 0, unit = "mm"),
    legend.spacing = unit(0, "mm"),
    strip.background = element_blank(),
    strip.text = element_blank()
  )
pdfname = "PAM-Desiccation-RWC-plot.pdf"
save_plot(pdfname, plot = p1, width = 80, height = 80*1.5)
```



::: {#fig-pam-desiccation}

```{r}
image_read(str_replace(pdfname, "pdf", "png"))
```

Non-linear response of the effective quantum yield of photosystem II (ΔFv/Fm') to relative water content across distinct hydration states. 
The panels illustrate the physiological trajectories during desiccation, 30-minute rehydration, and 1-day rehydration. 
Observations are plotted as points, with distinct shapes indicating individuals.
Lines represent the population-level expected marginal means derived from a Bayesian hierarchical generalized additive model (HGAM). 
Shaded regions denote the 95% highest density credible intervals. 
:::




```{r}
library(emmeans)
library(bayestestR)
```


```{r}
eout = emmeans(bout_full, ~nstate|x, at = list(x = seq(0, 1, by = 0.01)), epred = TRUE)

out = regrid(eout, transform = T) |> 
  contrast("pairwise") |> 
  describe_posterior(ci = 1, rope_range = c(-0.05, 0.05)) |> 
  as_tibble()
```


```{r}

out = out |> 
  mutate(important = (pd > 0.975)&(ROPE_Percentage < 0.025))

p1 = 
  ggplot(out) + 
  geom_hline(yintercept = 0.975) +
  geom_line(
    aes(
      x = x, y = pd, color = important, group = contrast
    )
  )
p2 = 
  ggplot(out) + 
  geom_hline(yintercept = 0.025) +
  geom_line(
    aes(
      x = x, y = ROPE_Percentage, color = important, group = contrast
    )
  )
p1 / p2
```




















