---
title: "2026 *Ulva australis* study"
title-block-banner: true
date: today
date-format: "YYYY MMM MM"
author: 
  - name: Dai Sumiyoshi
  - name: Kaiki Tajiri
  - name: Masahiro Suzuki
  - name: Gregory N. Nishihara
  - name: Ryuta Terada
execute: 
  echo: false
  message: false
  warning: false
---

The response of the photochemical efficiency (&Delta; F/Fm&prime;) in the green alga, 
*Ulva australis* from Nagashima Island, Kagoshima, Japan.
The experiments were conducted at temperature treatments of 4 to 36 &deg;C 
after 1, 3, 5, and 7 days of incubation at
50 &mu;mol photons m^-2^ s^-1^ and a 12L12D photoperiod.

```{r}
#| include: false
library(tidyverse)
library(ggpubr)
library(ggtext)
library(magick)
library(patchwork)
library(showtext)
library(brms)
library(tidybayes)
library(gt)
library(gnnlab)
library(utils)
library(readxl)
```

```{r}
#| include: false
font_add_google("Noto Sans", family = "ns")
theme_pubr(10, "ns", border = TRUE) |> 
  set_theme()
theme_pubr() |> 
  replace_theme(
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    axis.title.x = element_markdown(family = "ns", margin = margin(0.5, 0, 0.5, 0, "lines")),
    axis.title.y = element_markdown(family = "ns", angle= 90, margin = margin(0, 0.5, 0, 0.5, "lines")),
  )
showtext_auto()
folder = "~/Dropbox/MS_Terada_Greg/"
folder = "./Data/"
fnames = dir(folder, recursive = TRUE)
```


```{r}
ftext = "PAM_Temp"
fname = str_c(folder, str_subset(fnames, ftext)) 
esheets = excel_sheets(fname)
data = tibble(fname, esheets) |> 
  mutate(data = map2(fname, esheets, \(fname = fname, esheets = esheets) {
    read_xlsx(fname, sheet = esheets[1]) |> select(Temp, PAR, matches("Y\\(II\\)"))
  } )) |> 
  mutate(Days = str_extract(esheets, "[0-9]+")) |> 
  select(Days, data) |> 
  unnest(data) |> 
  pivot_longer(matches("^Y")) |> 
  mutate(ID = as.numeric(str_extract(name, "[0-9]+")))

```

```{r}
xlabel = "Temperature (&deg;C)"
ylabel = "Effective quantum yield (&Delta;F/Fm&prime;)"
xbreaks = unique(data$Temp)
xbreaks = seq(min(xbreaks), max(xbreaks), by = 4)
# 1. Create a label dataframe 
# We take distinct categories and generate the "A: ..." label here
labeldf = data |> expand(Days, x = min(Temp), y = 1) |> 
  mutate(
    l = str_glue("({LETTERS[seq_along(Days)]}) Incubation duration: {Days}-d ")
  )

p1 = ggplot(data) + 
  geom_point(aes(x = Temp, y = value), 
             position = position_jitter(width = 0.2), 
             shape = 21,
             fill = "black",
             stroke = 0,
             alpha = 0.8) +
  geom_richtext(aes(x = x, y = y, label = l),
                data = labeldf,
    vjust = 1, hjust = 0,
    color = "transparent",
    fill = "transparent",
    text.color = "black",
    family = "ns", 
    size = 3
  ) +
  scale_x_continuous(xlabel, breaks = xbreaks, limits = range(xbreaks)) +
  scale_y_continuous(ylabel, breaks = seq(0, 1, by = 0.25), limits = c(0, 1)) +
  facet_wrap(vars(Days), ncol = 2) +
  theme(
    strip.text = element_blank(),
    strip.background = element_blank()
  )

pdfname = "PAM-Temperature-plot.pdf"
save_plot(pdfname, plot = p1, width = 80 * 2, height = 80)
```



::: {#fig-pam-temp}


```{r}
image_read(str_replace(pdfname, "pdf", "png"))
```

The response of the photochemical efficiency (&Delta; F/Fm&prime;) in the green alga, 
*Ulva australis* from Nagashima Island, Kagoshima, Japan.
The experiments were conducted at temperature treatments of 4 to 36 &deg;C 
after 1, 3, 5, and 7 days of incubation at
50 &mu;mol photons m^-2^ s^-1^ and a 12L12D photoperiod.
The points are jittered horizontally for clarity (n = 10).

:::

