---
title: "2026 *Ulva australis* PAM-Temperature Inhibition Analysis"
title-block-banner: true
date: today
date-format: "YYYY MMM MM"
author: 
  - name: Dai Sumiyoshi
  - name: Kaiki Tajiri
  - name: Masahiro Suzuki
  - name: Gregory N. Nishihara
  - name: Ryuta Terada
execute: 
  echo: false
  message: false
  warning: false
---

```{r}
#| include: false
library(tidyverse)
library(ggpubr)
library(ggtext)
library(magick)
library(patchwork)
library(showtext)
library(brms)
library(tidybayes)
library(gt)
library(gnnlab)
library(utils)
library(readxl)
```

```{r}
#| include: false
font_add_google("Noto Sans", family = "ns")
theme_pubr(10, "ns", border = TRUE) |> 
  set_theme()
theme_pubr() |> 
  replace_theme(
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    axis.title.x = element_markdown(family = "ns", margin = margin(0.5, 0, 0.5, 0, "lines")),
    axis.title.y = element_markdown(family = "ns", angle= 90, margin = margin(0, 0.5, 0, 0.5, "lines")),
  )
showtext_auto()
folder = "~/Dropbox/MS_Terada_Greg/"
folder = "./Data/"
fnames = dir(folder, recursive = TRUE)
```


```{r}
#| include: false
ftext = "Inhibition"
fname = str_c(folder, str_subset(fnames, ftext)) 
esheets = excel_sheets(fname)
data = tibble(fname, esheets) |> 
  mutate(data = map2(fname, esheets, \(fname = fname, esheets = esheets) {
    read_xlsx(fname, sheet = esheets[1]) |> select(Temp, Hours, PAR, matches("Y\\(II\\)"))
  } ))

data = data |> 
  select(esheets, data) |> 
  unnest(data) |> 
  pivot_longer(matches("^Y")) |> 
  mutate(ID = as.numeric(str_extract(name, "[0-9]+"))) |> 
  mutate(Temp = factor(Temp, levels = c(28, 16, 4)))
```

```{r}
dynamic_labeller <- as_labeller(function(x) {
  str_glue("{x} &mu;mol photons m<sup>-2</sup> s<sup>-1</sup>")
})

xlabel = "Time of exposure (hours)"
ylabel = "Effective quantum yield (&Delta;F/Fm&prime;)"
xbreaks = unique(data$Hours)
xbreaks = seq(min(xbreaks), max(xbreaks), by = 2)

labeldf =
  data |> expand(Temp, PAR, x = min(Hours), y = 1) |> 
  mutate(
    l = str_glue("({LETTERS[seq_along(Temp)]}) {Temp}&deg;C")
  )
labeldf2 = 
  tibble(x = c(18-6)/2 + 6, y = 0) |> 
  mutate(l = "Dim-light acclimation")

p1 = ggplot(data) + 
  geom_rect(aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax), data = tibble(xmin = 6, xmax = 18, ymin = 0, ymax = 1), fill = "grey") +
  geom_richtext(aes(x = x, y = y, label = l),
                data = labeldf2,
    vjust = 0, hjust = 0.5,
    color = "transparent",
    fill = "transparent",
    text.color = "black",
    family = "ns", 
    size = 3
  ) +
  geom_point(aes(x = Hours, y = value), 
             position = position_jitter(width = 0.2), 
             shape = 21,
             fill = "black",
             stroke = 0,
             alpha = 0.8) +
  geom_richtext(aes(x = x, y = y, label = l),
                data = labeldf,
    vjust = 1, hjust = 0,
    color = "transparent",
    fill = "transparent",
    text.color = "black",
    family = "ns", 
    size = 3
  ) +
  scale_x_continuous(xlabel, breaks = xbreaks, limits = range(xbreaks)) +
  scale_y_continuous(ylabel, breaks = seq(0, 1, by = 0.25), limits = c(0, 1)) +
  facet_grid(rows = vars(Temp), cols = vars(PAR), labeller = labeller(PAR = dynamic_labeller)) +
  theme(
    strip.text.y = element_blank(),
    strip.text.x = element_markdown(family = "ns"),
    strip.background = element_blank()
  )
pdfname = "PAM-Temperature-Inhibition-plot.pdf"
save_plot(pdfname, plot = p1, width = 80 * 2, height = 80*1.5)
```



::: {#fig-pam-temp-inhib}


```{r}
image_read(str_replace(pdfname, "pdf", "png"))
```

The hourly response of the effective quantum yield (&Delta; F/Fm&prime;) in the green alga, 
*Ulva australis* from Nagashima Island, Kagoshima, Japan.
The experiments were conducted at temperature treatments of (A, B) 28&deg;C, (C, D) 16&deg;C, and (E, F) 4&deg;C crossed with 
irradiances of (A, C, D) 200 &mu;mol photons m^-2^ s^-1^ and (B, D, F) 1000 &mu;mol photons m^-2^ s^-1^.
The explants were acclimated for 12 hours of dim-light treatment (20 &mu;mol photons m^-2^ s^-1^) and exposed to 6 hours of light.
The points are jittered horizontally for clarity (n = 10).
:::

