---
title: "2026 *Ulva australis* PAM-Temperature Inhibition Analysis"
title-block-banner: true
date: today
date-format: "YYYY MMM MM"
author: 
  - name: Dai Sumiyoshi
  - name: Kaiki Tajiri
  - name: Masahiro Suzuki
  - name: Gregory N. Nishihara
  - name: Ryuta Terada
execute: 
  freeze: true
  cache: true
  echo: false
  message: false
  warning: false
editor: 
  markdown: 
    wrap: 72
bibliography: references.bib
format:
  html:
    grid:
      margin-width: 200px
      sidebar-width: 200px
      body-width: 900px
    tbl-cap-location: top
    css: custom.css
---

## Methods

We analyzed the effective quantum yields observed when *U. australis* was exposed to 200 &mu;mol photons m^-2^ s^-1^ and 1000 &mu;mol photons m^-2^ s^-1^ with separate models.
The response of the effective quantum yields were modeled with a linear model (Eq. @eq-model-form), where we examined the main effects of temperature ($\beta_t$) and time ($\beta_h$) and their interactions ($\beta_{t,h}$); the $\beta_0$ coefficient is the model intercept.

$$
logit(\mu) = \beta_0 + \beta_h + \beta_t + \beta_{h,t}
$${#eq-model-form}

We assigned a beta distribution for the low light exposure model and a zero-inflated beta distribution for the high light exposure model.
Beta distributions are bounded on the open continuous interval $[0, 1]$ and does not include exact zeros and ones.
Zero inflated beta distributions includes the presence of zeros as a separate Bernoulli model that we have used in the past [@borlongan2020;@terada2024]. 
The precision parameter ($\phi$) of both of these models was modeled as a linear model of temperature and on the log scale.
The zero-inflation rate of the high light exposure model was modeled as a linear combination of temperature, time, and their interaction on the logit scale.
Temperature and exposure time were defined as a factor variable, with three levels for temperature and three levels (0 hours, 6 hours, and 18 hours) for exposure time.

The model coefficients were assigned weakly informative distributions on the link scale.
A Gaussian distribution with a location of one and a scale of one was applied to the $\beta_0$ coefficient, 
and a Gaussian distribution with a location of zero and scale of one was applied to the remaining $\beta$ coefficients for Eq. {#eq-model-form}.
Similar priors were applied to the coefficients of the linear model for the zero-inflation rate.
A Gaussian distribution with a location of zero and a scale of two was assigned to the coefficients for the linear model for the precision.

```{r}
#| include: false
library(tidyverse)
library(ggpubr)
library(ggtext)
library(magick)
library(patchwork)
library(showtext)
library(future)
library(brms)
library(tidybayes)
library(bayestestR)
library(emmeans)
library(gt)
library(gnnlab)
library(utils)
library(readxl)
```

```{r}
font_add_google("Noto Sans", family = "ns")
theme_pubr(10, "ns", border = TRUE) |> 
  set_theme()
theme_pubr() |> 
  replace_theme(
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    axis.title.x = element_markdown(family = "ns", margin = margin(0.5, 0, 0.5, 0, "lines")),
    axis.title.y = element_markdown(family = "ns", angle= 90, margin = margin(0, 0.5, 0, 0.5, "lines")),
  )
showtext_auto()
folder = "~/Dropbox/MS_Terada_Greg/"
folder = "./Data/"
fnames = dir(folder, recursive = TRUE)
```

```{r}
#| include: false
ftext = "Inhibition"
fname = str_c(folder, str_subset(fnames, ftext)) 
esheets = excel_sheets(fname)
data = tibble(fname, esheets) |> 
  mutate(data = map2(fname, esheets, \(fname = fname, esheets = esheets) {
    read_xlsx(fname, sheet = esheets[1]) |> select(Temp, Hours, PAR, matches("Y\\(II\\)"))
  } ))

data = data |> 
  select(esheets, data) |> 
  unnest(data) |> 
  pivot_longer(matches("^Y")) |> 
  mutate(ID = as.numeric(str_extract(name, "[0-9]+"))) |> 
  mutate(Temp = factor(Temp, levels = c(28, 16, 4)))
```

```{r}
dynamic_labeller <- as_labeller(function(x) {
  str_glue("{x} &mu;mol photons m<sup>-2</sup> s<sup>-1</sup>")
})

xlabel = "Time of exposure (hours)"
ylabel = "Effective quantum yield (&Delta;F/Fm&prime;)"
xbreaks = unique(data$Hours)
xbreaks = seq(min(xbreaks), max(xbreaks), by = 2)

labeldf =
  data |> expand(Temp, PAR, x = min(Hours), y = 1) |> 
  mutate(
    l = str_glue("({LETTERS[seq_along(Temp)]}) {Temp}&deg;C")
  )
labeldf2 = 
  tibble(x = c(18-6)/2 + 6, y = 0) |> 
  mutate(l = "Dim-light acclimation")

p1 = ggplot(data) + 
  geom_rect(aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax), data = tibble(xmin = 6, xmax = 18, ymin = 0, ymax = 1), fill = "grey") +
  geom_richtext(aes(x = x, y = y, label = l),
                data = labeldf2,
                vjust = 0, hjust = 0.5,
                color = "transparent",
                fill = "transparent",
                text.color = "black",
                family = "ns", 
                size = 3
  ) +
  geom_point(aes(x = Hours, y = value), 
             position = position_jitter(width = 0.2), 
             shape = 21,
             fill = "black",
             stroke = 0,
             alpha = 0.8) +
  geom_richtext(aes(x = x, y = y, label = l),
                data = labeldf,
                vjust = 1, hjust = 0,
                color = "transparent",
                fill = "transparent",
                text.color = "black",
                family = "ns", 
                size = 3
  ) +
  scale_x_continuous(xlabel, breaks = xbreaks, limits = range(xbreaks)) +
  scale_y_continuous(ylabel, breaks = seq(0, 1, by = 0.25), limits = c(0, 1)) +
  facet_grid(rows = vars(Temp), cols = vars(PAR), labeller = labeller(PAR = dynamic_labeller)) +
  theme(
    strip.text.y = element_blank(),
    strip.text.x = element_markdown(family = "ns"),
    strip.background = element_blank()
  )
pdfname = here::here("Plots/PAM-Temperature-Inhibition-plot.pdf")
save_plot(pdfname, plot = p1, width = 80 * 2, height = 80*1.5)
```


## Results 

::: {#fig-observations}

```{r}
image_read(str_replace(pdfname, "pdf", "png"))
```

The hourly response of the effective quantum yield (&Delta; F/Fm&prime;) in the
green alga, *Ulva australis* from Nagashima Island, Kagoshima, Japan.
The experiments were conducted at temperature treatments of (A, B) 28&deg;C,
(C, D) 16&deg;C, and (E, F) 4&deg;C crossed with light treatments of (A, C, D) 200
&mu;mol photons m^-2^ s^-1^ and (B, D, F) 1000 &mu;mol photons m^-2^ s^-1^.
The explants were exposed to 6 hours of light followed by 12 hours of dim-light treatment (20
&mu;mol photons m^-2^ s^-1^).
The points are jittered horizontally for clarity (n = 10).

:::


::: {#tbl-observation-summary}

```{r}
data |> 
  arrange(Temp, PAR) |> 
  group_by(Temp, PAR, Hours) |> 
  reframe(across(value, list(m = mean, sd = sd, se = se, n = length))) |> 
  gt() |> 
  cols_label(
    Temp = "Temperature",
    PAR = "Irradiance",
    Hours = "Time of exposure",
    value_m = "Mean",
    value_sd = "Std. Dev.",
    value_se = "Std. Error",
    value_n = "N"
  ) |> 
  fmt_number(columns = c(value_m, value_sd), decimals = 3) |> 
  fmt_number(columns = value_se, decimals = 3) |> 
  cols_align(align = "right") |> 
  tab_footnote(
    footnote = html("&mu;mol photons m<sup>-2</sup> s<sup>-1</sup>"),
    locations = cells_column_labels(columns = PAR)
  ) |> 
  tab_footnote(
    footnote = html("&deg;C"),
    locations = cells_column_labels(columns = Temp)
  ) |> 
  tab_footnote(
    footnote = "hours",
    locations = cells_column_labels(columns = Hours),
  ) |> 
  opt_footnote_spec(spec_ref = "^xb", spec_ftr = "(x)") |> 
    tab_style(
    style = cell_text(align = "left"),
    locations = cells_footnotes()
  )
```

The mean, standard deviation (Std. Dev.), and standard error (Std. Err.) of the effective quantum yield (&Delta;Fv/Fm&prime;).

:::

```{r}
data0200 = data |> 
  filter(PAR != 1000) |> 
  filter(Hours %in% c(0, 6, 18)) |> 
  mutate(
    Temp = factor(Temp, levels = c("4", "16", "28")),
    Hours = factor(Hours, levels = c(0, 6, 18))
    )

data1000 = data |> 
  filter(PAR == 1000) |> 
  filter(Hours %in% c(0, 6, 18)) |> 
  mutate(
    Temp = factor(Temp, levels = c("4", "16", "28")),
    Hours = factor(Hours, levels = c(0, 6, 18))
    )
```

```{r}
# Important note: When a factor only has two levels,
# it is very difficult to use it as a random effect!

model0200 = bf(
  value  ~ Temp * Hours,
  phi ~ Temp
  ) + Beta()

model1000 = bf(
  value  ~ Temp * Hours,
  zi     ~ Temp * Hours, 
  phi ~ Temp
  ) + 
  zero_inflated_beta()

priors0200 = c(
  prior(normal(1, 1),      class = Intercept),
  prior(normal(0, 1),      class = b),
  prior(normal(1, 1),      dpar = phi, class = Intercept),
  prior(normal(0, 2),      dpar = phi, class = b)
)
priors1000 = c(
  prior(normal(1, 1),      class = Intercept),
  prior(normal(0, 1),      class = b),
  prior(normal(1, 1),      class = Intercept,    dpar = zi),
  prior(normal(0, 1),      class = b,         　 dpar = zi),
  prior(normal(1, 1),      dpar = phi, class = Intercept),
  prior(normal(0, 2),      dpar = phi, class = b)
)

cores = chains = 5
seed = 2026
warmup = 4000
iterations = warmup + 10000
# file_refit = "never"
control = list(adapt_delta = 0.9, max_treedepth = 20)
filename0200 = "pam-inhibition-full-model0200"
filename1000 = "pam-inhibition-full-model1000"

# file.remove(str_c(filename0200, ".rds"))
# file.exists(str_c(filename0200, ".rds"))
# file.remove(str_c(filename1000, ".rds"))
# file.exists(str_c(filename1000, ".rds"))
# 
# workers = "mimir"
# 
# if(!file.exists(str_c(filename0200, ".rds"))) {
#   plan(cluster, workers = workers)
#   futurefull0200 = 
#     future({
#       brm(
#         model0200,
#         data = data0200,
#         prior = priors0200,
#         chains = chains,
#         cores = cores,
#         seed = seed,
#         iter = iterations,
#         warmup = warmup,
#         refresh = 0,
#         backend = "cmdstanr",
#         control = control,
#         save_pars = save_pars(all = TRUE)
#       )
#     }, seed = seed)
# }
# 
# if(!file.exists(str_c(filename1000, ".rds"))) {
#   plan(cluster, workers = workers)
#   futurefull1000 = 
#     future({
#       brm(
#         model1000,
#         data = data1000,
#         prior = priors1000,
#         chains = chains,
#         cores = cores,
#         seed = seed,
#         iter = iterations,
#         warmup = warmup,
#         refresh = 0,
#         backend = "cmdstanr",
#         control = control,
#         save_pars = save_pars(all = TRUE)
#       )
#     }, seed = seed)
# }
# 
# while(! (resolved(futurefull0200) & resolved(futurefull1000))) {
#   message("Sampling in progress.")
#   Sys.sleep(1)
# }

# bout0200 = value(futurefull0200)
# bout0200 |> write_rds(str_c(filename0200, ".rds"))
# bout1000 = value(futurefull1000)
# bout1000 |> write_rds(str_c(filename1000, ".rds"))

bout0200 = read_rds(str_c(filename0200, ".rds"))
bout1000 = read_rds(str_c(filename1000, ".rds"))
# plan(sequential); gc()

pdata0200_df = data0200 |> expand(Temp, Hours) 
pdata1000_df = data1000 |> expand(Temp, Hours) 
pdata0200_df = pdata0200_df |> add_epred_draws(bout0200, allow_new_levels = TRUE) |> mean_hdci(.epred)
pdata1000_df = pdata1000_df |> add_epred_draws(bout1000, allow_new_levels = TRUE) |> mean_hdci(.epred)

p0200_1 = pp_check(bout0200, type = "dens_overlay_grouped",          group = "Temp",                     ndraws = 50) + facet_wrap(vars(group), nrow = 1)
p0200_2 = pp_check(bout0200, type = "stat_grouped", stat = "median", group = "Temp",   binwidth = 0.001, ndraws = 500) + facet_wrap(vars(group), nrow = 1)
p0200_3 = pp_check(bout0200, type = "stat_grouped", stat = "mad",    group = "Temp",   binwidth = 0.001, ndraws = 500) + facet_wrap(vars(group), nrow = 1)
p0200_4 = pp_check(bout0200, type = "stat_grouped", stat = "mean",   group = "Temp",   binwidth = 0.001, ndraws = 500) + facet_wrap(vars(group), nrow = 1)
p0200_5 = pp_check(bout0200, type = "stat_grouped", stat = "sd",     group = "Temp",   binwidth = 0.001, ndraws = 500) + facet_wrap(vars(group), nrow = 1)
p1000_1 = pp_check(bout1000, type = "dens_overlay_grouped",          group = "Temp",                     ndraws = 50) + facet_wrap(vars(group), nrow = 1)
p1000_2 = pp_check(bout1000, type = "stat_grouped", stat = "median", group = "Temp",   binwidth = 0.001, ndraws = 500) + facet_wrap(vars(group), nrow = 1)
p1000_3 = pp_check(bout1000, type = "stat_grouped", stat = "mad",    group = "Temp",   binwidth = 0.001, ndraws = 500) + facet_wrap(vars(group), nrow = 1)
p1000_4 = pp_check(bout1000, type = "stat_grouped", stat = "mean",   group = "Temp",   binwidth = 0.001, ndraws = 500) + facet_wrap(vars(group), nrow = 1)
p1000_5 = pp_check(bout1000, type = "stat_grouped", stat = "sd",     group = "Temp",   binwidth = 0.001, ndraws = 500) + facet_wrap(vars(group), nrow = 1)


design = "
AF
BG
CH
DI
EJ"

pout = p0200_1 + p0200_2 + p0200_3 + p0200_4 + p0200_5 + p1000_1 + p1000_2 + p1000_3 + p1000_4 + p1000_5 + plot_layout(design = design)

pdfname = here::here("Plots/pam-inhibition-diagnostics.pdf")
pngname = str_replace(pdfname, "pdf", "png")
ggsave(pdfname, plot = pout, width = 6*160, height = 3*100, units = "mm")
image_read_pdf(pdfname) |> image_write(pngname)
```

```{r}
tab0200 = 
  emmeans(bout0200, ~Hours|Temp) |> contrast("trt.vs.ctrl1") |> describe_posterior(rope_ci = 1, rope_range = c(-0.1, 0.1) * pi/sqrt(3)) |> 
  as_tibble()
zip1000 = 
  emmeans(bout1000, ~Hours|Temp, dpar = "zi") |> contrast("trt.vs.ctrl1") |> describe_posterior(rope_ci = 1, rope_range = c(-0.1, 0.1) * pi/sqrt(3)) |> 
    as_tibble()
tab1000 = 
  emmeans(bout1000, ~Hours|Temp) |> contrast("trt.vs.ctrl1") |> describe_posterior(rope_ci = 1, rope_range = c(-0.1, 0.1) * pi/sqrt(3)) |> 
  as_tibble()

tab1000 = bind_rows(
tab1000 |> filter(!str_detect(Temp, "4")),
zip1000 |> filter(str_detect(Temp, "4"))
) |> 
  arrange(Temp)

```



::: {#tbl-multiple-comparisons-0200}

```{r}
tab0200 |> 
  mutate(contrast  = str_replace(contrast, "Hours0", "0 hrs")) |> 
  mutate(contrast  = str_replace(contrast, "Hours6", "6 hrs")) |> 
  mutate(contrast  = str_replace(contrast, "Hours18", "18 hrs")) |> 
  mutate(hdci = sprintf("%0.3f to %0.3f", CI_low, CI_high)) |> 
  select(Temp, contrast, Median, hdci, pd, ROPE_Percentage) |> 
  gt() |>
  cols_label(
    Temp = html("Temperature"),
    contrast = "Contrast",
    Median = "Median effect",
    hdci = "95% HDCI",
    pd = "PD",
    ROPE_Percentage = "Pr. ROPE"
  ) |> 
  fmt_number(columns = c(Median), decimals = 3) |> 
  fmt_number(columns = c(pd), decimals = 1, scale_by = 100, pattern = "{x}%") |> 
  fmt_number(columns = c(ROPE_Percentage), decimals = 1, scale_by = 100, pattern = "{x}%") |> 
  cols_align(align = "right") |> 
  tab_footnote(
    footnote = html("&deg;C"),
    locations = cells_column_labels(columns = Temp)
  ) |> 
  tab_footnote(
    footnote = "95% highest density credible interval",
    locations = cells_column_labels(columns = hdci)
  ) |> 
  tab_footnote(
    footnote = "Probability of direction",
    locations = cells_column_labels(columns = pd)
  ) |> 
  tab_footnote(
    footnote = "Probability of being within the region of practical equivalence",
    locations = cells_column_labels(columns = ROPE_Percentage)
  ) |> 
  tab_footnote(
    footnote = "On the logit scale",
    locations = cells_column_labels(columns = c(Median, hdci)),
    placement = "left"
  ) |> 
  opt_footnote_spec(spec_ref = "^xb", spec_ftr = "(x)") |> 
    tab_style(
    style = cell_text(align = "left"),
    locations = cells_footnotes()
  )
```

Comparing the effect of time on the effective quantum yields in the 200 &mu;mol photons m^-2^ s^-1^ exposure experiment. 
The median effect and the 95% highest density credible interval (HDCI) is on the logit scale. 
PD is the Probability of Direction, which is an index of the proportion of the posterior distribution that is in the direction of the sign of the median effect and ranges for 0.5 to 1.
Pr. ROPE is the probability that 100% of the posterior distribution falls within the Region of Practical Equivalence (ROPE).
ROPE ranges from 0 to 1, and if the Pr. ROPE is above 97.5% than there is a high likelihood that the posterior distribution of the effect falls completely within the ROPE.
If the Pr. ROPE is below 2.5% then there is a high likelihood that the posterior distribution of the effect is not in side the ROPE.
the 100% HDCI is within the region of practical equivalence (ROPE) and ranges from 0 to 1.

:::

The response of effective quantum yield exposed to low light across different temperatures was analyzed and the posterior contrasts were calculated by comparing performance at 6 and 18 hours against the effective quantum yield at 0 hours (baseline). 
Estimates are reported on the logit scale (@tbl-multiple-comparisons-0200). 
To determine whether these changes were biologically meaningful, we calculated the Probability of Direction (PD) and the percentage of the posterior distribution falling within a Region of Practical Equivalence (ROPE), defined a priori as $\pm 0.10 \times \pi/\sqrt{3}$ [@Kruschke2013;@kruschke2021;@ortega2017@kruschke2017].

We detected sustained depression of the effective quantum yield in the 4&deg;C treatment (@tbl-multiple-comparisons-0200).
Exposure to low temperature induced a decrease in effective quantum yield (@fig-observations) and 
by 6 hours, the effective quantum yield dropped substantially from baseline.
The probability of direction (PD) was 100% for this decrease, with a 0.0% probability that the posterior distribution fell within the ROPE, indicating a significant reduction in effective quantum yield. 
After dim-light acclimation, the effective quantum yield recovered, however the yield remained significantly depressed compared to the baseline,
reflecting sustained physiological stress without full recovery.

During the 16&deg;C exposure experiment, the effective quantum yield showed a small but significant reduction in yield at 6 hours (@tbl-multiple-comparisons-0200). 
However, after dim-light treatment, the effective quantum yield completely recovered. 
The median difference from baseline approached zero and 99.5% of the posterior distribution fell completely within the ROPE. 
This provides strong evidence that 12 hours of dim-light treatment returned the effective quantum yield to their baseline performance.

At 28 &deg;C, the results were similar to that of the 16&deg;C exposure experiment, and was characterized by a significant drop at 6 hours (@tbl-multiple-comparisons-0200). 
The effective quantum yield recovered to slightly above baseline levels after the dim-light treatment.
Despite the confidence in the direction of this increase, 63.4% of the posterior distribution fell within the ROPE.


::: {#tbl-multiple-comparisons-1000}

```{r}
tab1000 |> 
  mutate(contrast  = str_replace(contrast, "Hours0", "0 hrs")) |> 
  mutate(contrast  = str_replace(contrast, "Hours6", "6 hrs")) |> 
  mutate(contrast  = str_replace(contrast, "Hours18", "18 hrs")) |> 
  mutate(hdci = sprintf("%0.3f to %0.3f", CI_low, CI_high)) |> 
  select(Temp, contrast, Median, hdci, pd, ROPE_Percentage) |> 
  gt() |>
  cols_label(
    Temp = html("Temperature"),
    contrast = "Contrast",
    Median = "Median effect",
    hdci = "95% HDCI",
    pd = "PD",
    ROPE_Percentage = "Pr. ROPE"
  ) |> 
  fmt_number(columns = c(Median), decimals = 3) |> 
  fmt_number(columns = c(pd), decimals = 1, scale_by = 100, pattern = "{x}%") |> 
  fmt_number(columns = c(ROPE_Percentage), decimals = 1, scale_by = 100, pattern = "{x}%") |> 
  cols_align(align = "right") |> 
  tab_footnote(
    footnote = html("&deg;C"),
    locations = cells_column_labels(columns = Temp)
  ) |> 
  tab_footnote(
    footnote = "95% highest density credible interval",
    locations = cells_column_labels(columns = hdci)
  ) |> 
  tab_footnote(
    footnote = "Probability of direction",
    locations = cells_column_labels(columns = pd)
  ) |> 
  tab_footnote(
    footnote = "Probability of being within the region of practical equivalence",
    locations = cells_column_labels(columns = ROPE_Percentage)
  ) |> 
  tab_footnote(
    footnote = "On the logit scale",
    locations = cells_column_labels(columns = c(Median, hdci)),
    placement = "left"
  ) |> 
  opt_footnote_spec(spec_ref = "^xb", spec_ftr = "(x)") |> 
  tab_style(
    style = cell_text(align = "left"),
    locations = cells_footnotes()
  )
```

Comparing the effect of time on the effective quantum yields in the 1000 &mu;mol photons m^-2^ s^-1^ exposure experiment at 16&deg;C and 28&deg;C.
Note that the comparisons done at 4&deg;C examine the increase in probability that the effective quantum yield will be zero.
The median effect and the 95% highest density credible interval (HDCI) is on the logit scale. 
PD is the Probability of Direction, which is an index of the proportion of the posterior distribution that is in the direction of the sign of the median effect and ranges for 0.5 to 1.
Pr. ROPE is the probability that 100% of the posterior distribution falls within the Region of Practical Equivalence (ROPE).
ROPE ranges from 0 to 1, and if the Pr. ROPE is above 97.5% than there is a high likelihood that the posterior distribution of the effect falls completely within the ROPE.
If the Pr. ROPE is below 2.5% then there is a high likelihood that the posterior distribution of the effect is not in side the ROPE.
the 100% HDCI is within the region of practical equivalence (ROPE) and ranges from 0 to 1.


:::

We also assesed the impact of high light exposure (1000 &mu;mol photons m^-2^ s^-1^) on the effective quantum yield across three temperature treatments; the posterior contrasts were evaluated comparing performance at 6 and 18 hours against the baseline (@fig-observations). 
As with the low light exposure experiment, the estimates of effective quantum yield are reported on the logit scale, and practical significance was assessed using a Probability of Direction (pd) and a Region of Practical Equivalence (ROPE) [@Kruschke2013;@kruschke2021;@ortega2017@kruschke2017].

At 4&deg;C, the high light exposure completely drove the effective quantum yields to zero after 6 hours and there was no recovery after dim-light treatment (@tbl-multiple-comparisons-1000).
The absence of non-zero observations cannot inform the median effect size under these conditions, hence their posterior estimates reflect the 
prior distributions assigned to the model and cannot meaningfully be compared.
Therefore, we compare the zero-inflation probability of the model.
Positive estimates on the logit scale indicate an increased probability of an effective quantum yield of zero, compared to the baseline.
After 6 hours of exposure, there was a significant increase in the probability of complete cessation of photosynthetic activity.
This risk of total failure is highly robust, characterized by a 99.9% Probability of Direction and only 0.2% of the posterior distribution falling within the ROPE. 
This state of severe impairment was sustained through the dim-light acclimation period, where the probability of zero yield remained definitively elevated above baseline.
Collectively, these zero-inflation contrasts confirm that the combined stress of low temperature and high light induces an immediate, severe, and sustained state of complete photo-inhibition.

Explants exposed to 1000 &mu;mol photons m^-2^ s^-1^ at 16&deg;C exhibited a severe reduction in quantum yield at 6 hours
and this drop was significant (@tbl-multiple-comparisons-1000). 
Despite the severity of the initial high-light exposure, performance demonstrated a complete recovery after treatment with dim-light.
The median difference returned to near-zero, with a PD of 96.6% and 100% of the posterior distribution fell entirely within the ROPE, indicating definitive practical equivalence with baseline performance.

At 28&deg;C, the 6-hour exposure resulted in a significant impairment in effective quantum yield (@tbl-multiple-comparisons-1000).
Unlike the full recovery observed at 16&deg;C, the recovery at 28&deg;C after dim-light acclimation was incomplete. 
The effective quantum yield remained slightly depressed compared to baseline.
The practical significance of this residual depression is ambiguous, as 44.7% of the posterior distribution remained within the ROPE.

To validate the appropriateness and predictive accuracy of the fitted Bayesian models, comprehensive graphical posterior predictive checks (PPCs) were conducted. 
Model fit was evaluated by comparing the observed experimental data to simulated datasets drawn from the joint posterior predictive distribution of the models. 
Density overlays confirmed that both the low-light model and the high-light model accurately replicated the overall shape and boundaries of the observed effective quantum yield distributions. 
Furthermore, summary-statistic PPCs demonstrated that the models reliably captured both the central tendencies (mean and median) and the dispersion (standard deviation and median absolute deviation) of the data across all temperature treatments. 
Crucially, the diagnostic checks verified that explicitly modeling the precision parameter ($\phi$) as a function of temperature successfully accounted for observed heteroscedasticity. 
Likewise, the zero-inflated architecture appropriately and accurately reproduced the total structural collapse (yields of absolute zero) observed under the 4&deg;C high-light treatment. 
These diagnostics confirm that the selected model structures provide robust, unbiased, and biologically realistic representations of the physiological data.

--------------------------------------------------------------------------------

## PD と Pr. ROPE について

Some papers that discuss the concept of the region of practical equivalence (ROPE) and the
probability of direction (PD) are @kruschke2021, @anderson2000, @ortega2017, @kruschke2015,
@kruschke2017, and @Kruschke2013.

**Probability of Direction (PD)** と **Probability of ROPE** の概念です。

### 1. Probability of Direction (PD): 方向性の確率

**「効果がプラス（またはマイナス）である確率は何％か？」**

* **定義:** 事後分布（推定されたパラメータの分布）のうち、符号が「中央値と同じ方向」にある割合です。つまり、効果が「正」または「負」であるという確信度を表します。
* **P値との関係:**
頻度論の P値と非常に強い相関があります。P値が「帰無仮説が正しいとした場合に極端なデータが得られる確率」であるのに対し、PDは「効果が存在する（ゼロではない）確率」に近いです。
* **PD  97.5%** は、両側検定の **P値  0.05** に相当します。
* **PD  99.5%** は、**P値  0.01** に相当します。
* PDが **1.00 (100%)** に近いほど、その効果が偶然ではなく、確実に特定の方向（増加または減少）に働いていることを示します。

### 2. Probability of ROPE: 実質的等価領域の確率

**「その効果は、実質的に『ゼロ』とみなせるか？」**

* **ROPE (Region of Practical Equivalence) とは:**
「効果が完全にゼロ」であることは稀ですが、「科学的・実用的に意味がないほど小さい（ほぼゼロ）」範囲は存在します。この範囲（例: -0.1 から 0.1 の間）を **ROPE** と呼びます。
* **確率の意味:**
事後分布の何％がこの ROPE の範囲内に収まっているかを示します。
* **解釈:**
- **Pr. ROPE が高い（例: > 0.95）:** 推定された効果のほとんどが「意味のない範囲」に入っています。「**実質的に差はない（帰無仮説を支持する）**」という結論を積極的に導くことができます。（P値では「差がない」ことを証明するのは難しいですが、ROPEなら可能です）
-  **Pr. ROPE が低い（例: < 0.025）:** 効果の大部分が「意味のない範囲」の外にあります。「**実質的な意味を持つ差がある**」可能性が高いことを示唆します。


### P値を用いた頻度論的解釈との対比まとめ

| 指標 | 問いかけ | 頻度論（P値）での類似概念 |
| --- | --- | --- |
| **PD (Probability of Direction)** | 「効果がある（ゼロではない）と言えるか？」 | **有意差検定 (NHST)**<br><br>PDが高いほど、P値は低い（有意）。 |
| **Pr. ROPE** | 「その効果は実質的に意味がある大きさか？」 | **同等性検定 (Equivalence Testing)**<br><br>サンプル数が多すぎて「統計的には有意だが、実用上は意味がない微差」が生じた場合、P値は小さくなりますが、Pr. ROPEは高くなる（差がないと判断できる）ことがあります。 |

### 論文の結果解釈への応用

* **PD = 1.000, Pr. ROPE < 0.001** の場合：
「効果は確実に存在し（PD = 1.00）、かつその大きさは実質的に意味がある（ROPE確率がほぼ 0）。」→ 非常に強い有意な効果。
* **PD = 0.992, Pr. ROPE = 0.992** の場合：
「方向性は一定しているように見えるが（PDが高い）、効果の大きさ自体は極めて小さく、実質的には『差がない』とみなせる（ROPE確率が高い）。」→ 統計的に検出されたとしても、科学的に考察できる結果ではない。


## 帰無仮説検定との比較

NHST（帰無仮説検定）に対する、PD（Probability of Direction）および ROPE（実質的等価領域）の主な優位性は、
**「『効果がない（帰無仮説）』ことを積極的に支持できるかどうか」**、
そして**「実質的な意味（効果の大きさ）」を評価できるか**という点にあります。

### 1. 「証拠がない」のか「無いという証拠」なのか？（NHSTの限界）

NHST（P値）の最大の問題点は、有意差が出なかった場合（）の解釈が曖昧なことです。

* **NHSTの場合:** 「帰無仮説を棄却できない（Fail to reject）」としか言えません。
- これは、「効果が本当にゼロである」のか、単に「データ不足（パワー不足）で検出できなかっただけ」なのかを区別できません。
- つまり、**「証拠がない（Absence of evidence）」**ことしか示せず、**「無いという証拠（Evidence of absence）」**にはなりません。
* **ROPEの場合:** 「実質的にゼロとみなせる範囲」を定義し、事後分布の大部分がそこに含まれれば、**「効果は実質的にゼロである」と積極的に結論づける（帰無仮説を受容する）**ことができます。
- これにより、「データは十分にあるが、効果は無視できるほど小さい」という強力な結論を導くことが可能です。



### 2. 「統計的に有意」vs.「実質的に意味がある」（サンプルの罠）

サンプルサイズ（N）が非常に大きい場合、NHSTには別の問題が発生します。

* **NHSTの場合:**  N が巨大であれば、極めて微細でどうでもいい差であっても、P値は小さくなり「統計的に有意」と判定されてしまいます。これは科学的なミスリードを生む原因となります。
* **PD & ROPEの場合:**
- **PD:** 効果の存在確率は高く出ますが（「確かにプラス方向である」）、それだけでは判断しません。
- **ROPE:** 効果の大きさが ROPE（実質的等価領域）の中に留まっていれば、いくら方向が確実（高PD）でも、**「実質的には意味がない（差はない）」**と判断できます。これにより、過剰な検定力による「意味のない有意差」をフィルタリングできます。



### 3. 直感的な理解（定義の明確さ）

* **P値（NHST）:** 「『もし帰無仮説が正しいとしたら』、今回のようなデータが得られる確率はどれくらいか？」という、条件付き確率の非常に回りくどい概念です。
* **PD:** 「効果がプラス（またはマイナス）である確率は何％か？」という、私たちが本当に知りたい直接的な確率です。

### まとめ

| 特徴 | NHST (P値) | ベイズ流 (PD & ROPE) |
| --- | --- | --- |
| **帰無仮説の扱い** | 棄却するか、保留するかのみ。「差がない」ことの証明は不可（悪魔の証明）。 | 「実質的に差がない」ことを確率的に証明（受容）できる。 |
| **有意差が出ない時** | 「データ不足」か「効果なし」か区別不能。 | ROPEを見ることで、「精度不足（不確実）」か「確実に効果なし」かを区別可能。 |
| **サンプル数過多** | 微細な差でも「有意」になり、ミスリードしやすい。 | 効果が微細なら ROPE に含まれるため、「実質的な差なし」と正しく判断できる。 |
| **解釈の直感性** | 誤解されやすい（P値を効果の確率と思いがち）。 | 直感的（PD = 効果が存在する確率）。 |

**結論として：**
論文の Results や Discussion において、「有意差が出なかった」ことを単に報告するのではなく、ROPE を用いて「この結果は、効果が科学的に無視できるレベルであることを示唆している（Accepting the Null）」と論じることができる点が、この手法の最大の強みです。

--------------------------------------------------------------------------------

## Model diagnostic plots


:::{#fig-overlay}
```{r}
p0200_1 + p1000_1 + plot_layout(ncol = 1) + plot_annotation(tag_levels = 'A')
```

Bayesian Posterior Predictive Check (PPC) Density Overlays for Effective Quantum Yield Models.

The plots represent a visual assessment of model fit (density overlays) comparing the observed distribution of effective quantum yield data to the data predicted by the fitted Bayesian models. 
The top row (A) displays results from the low-light model, while the bottom row (B) displays results from the high-light model. 
Individual panels within each row correspond to the temperature treatments: 4&deg;C, 16&deg;C, and 28&deg;C. 
In all panels, the thick dark blue line ($y$) represents the kernel density estimation of the observed experimental data. 
The multiple thin light blue lines ($y_{\text{rep}}$) represent density estimations calculated from 50 independent datasets simulated (replicated) from the posterior predictive distributions of the model. 
Row B captures the characteristics of the zero-inflated model, specifically showing the pronounced spike at exactly zero yield observed in the 4&deg;C panel due to total photoinhibition. 
High congruence between the thick and thin lines signifies an adequate model fit.
:::


:::{#fig-median}
```{r}
p0200_2 + p1000_2 + plot_layout(ncol = 1) + plot_annotation(tag_levels = 'A')

```

Posterior Predictive Check (PPC) of Median Effective Quantum Yield.
This plots illustrates a summary-statistic-based posterior predictive check, evaluating the ability of the model to reproduce the central tendency (median) of the observed data. 
The top row (A) displays the performance of the low-light model, and the bottom row (B) displays the high-light zero-inflated model. 
Columns indicate the temperature treatments: 4&deg;C, 16&deg;C, and 28&deg;C. 
In each panel, the solid dark blue vertical line represents the median calculated directly from the observed experimental data, $T(y)$. 
The light blue histogram displays the distribution of medians, $T(y_{\text{rep}})$, calculated from simulated datasets drawn from the posterior predictive distribution of the model. 
The consistent alignment of the observed medians within the main mass of the simulated distributions across treatments indicates that both models adequately capture the central tendency of the quantum yield responses.
Notably, in row B at 4&deg;C, the model accurately replicates the observed median of absolute zero, driven by the zero-inflation component.
:::

:::{#fig-mad}
```{r}
p0200_3 + p1000_3 + plot_layout(ncol = 1) + plot_annotation(tag_levels = 'A')

```

Posterior Predictive Check (PPC) of Median Absolute Deviation.
This plots illustrates a summary-statistic-based posterior predictive check, evaluating the ability of the model to capture the dispersion of the observed data by using the median absolute deviation (MAD). 
The top row (A) displays the performance of the low-light model, and the bottom row (B) displays the high-light zero-inflated model. 
Columns indicate the temperature treatments: 4&deg;C, 16&deg;C, and 28&deg;C. 
In each panel, the solid dark blue vertical line represents the MAD calculated directly from the observed experimental data, $T(y)$. 
The light blue histogram displays the distribution of MAD, $T(y_{\text{rep}})$, calculated from simulated datasets drawn from the posterior predictive distribution of the model. 
The alignment of the observed MAD values within the main mass of the simulated distributions across treatments indicates that the models adequately capture the variance present in the physiological responses. 
Notably, in row B at 4&deg;C, the observed MAD of absolute zero, resulting from complete photoinhibition, is accurately reproduced by the zero-inflated model architecture.
The zero-inflation rate drives the large variations observed, since there is a large spike at zero.
:::

:::{#fig-mean}
```{r}
p0200_4 + p1000_4 + plot_layout(ncol = 1) + plot_annotation(tag_levels = 'A')

```

Posterior Predictive Check (PPC) of Mean Effective Quantum Yield.
This plots illustrates a summary-statistic-based posterior predictive check, evaluating the ability of the model to reproduce the expected value (mean) of the observed data. 
The top row (A) displays the performance of the low-light model, and the bottom row (B) displays the high-light zero-inflated model. 
Columns indicate the temperature treatments: 4&deg;C, 16&deg;C, and 28&deg;C. 
In each panel, the solid dark blue vertical line represents the mean calculated directly from the observed experimental data, $T(y)$. 
The light blue histogram displays the distribution of means, $T(y_{\text{rep}})$, calculated from simulated datasets drawn from the posterior predictive distribution of the model. 
The consistent alignment of the observed means within the main mass of the simulated distributions across all treatments indicates that both models adequately capture the overall expected values of the physiological responses. 
The wider dispersion of the simulated means in row B, particularly under the 4&deg;C treatment, visually reflects the increased predictive variance inherent in the zero-inflated mixture model, as the simulated presence or absence of structural zeros trial-to-trial heavily influences the calculated mean of each replicated dataset.
:::

:::{#fig-sd}
```{r}
p0200_5 + p1000_5 + plot_layout(ncol = 1) + plot_annotation(tag_levels = 'A')

```

Posterior Predictive Check (PPC) of Standard Deviation.
This plots illustrates a summary-statistic-based posterior predictive check, evaluating the ability of the model to reproduce the standard deviation (sd) of the observed data. 
The top row (A) displays the performance of the low-light model, and the bottom row (B) displays the high-light zero-inflated model. 
Columns indicate the temperature treatments: 4&deg;C, 16&deg;C, and 28&deg;C. 
In each panel, the solid dark blue vertical line represents the standard deviation calculated directly from the observed experimental data, $T(y)$. 
The light blue histogram displays the distribution of standard deviations, $T(y_{\text{rep}})$, calculated from simulated datasets drawn from the posterior predictive distribution of the model. 
The consistent alignment of the observed standard deviations within the main mass of the simulated distributions across all treatments indicates that both models adequately capture the variance of the physiological responses. 
Similar to the mean summary statistic, the broader dispersion of the simulated standard deviations in row B reflects the additional trial-to-trial variance naturally introduced by the zero-inflated mixture modeling process.
:::



