---
title: "2026 *Ulva australis* PAM Kinetics"
title-block-banner: true
date: today
date-format: "YYYY MMM MM"
author: 
  - name: Dai Sumiyoshi
  - name: Kaiki Tajiri
  - name: Masahiro Suzuki
  - name: Gregory N. Nishihara
  - name: Ryuta Terada
execute: 
  freeze: true
  cache: true
  echo: false
  message: false
  warning: false
editor: 
  markdown: 
    wrap: 72
bibliography: references.bib
format:
  html:
    grid:
      margin-width: 200px
      sidebar-width: 200px
      body-width: 900px
    tbl-cap-location: top
    css: custom.css
---

```{r}
#| include: false
library(tidyverse)
library(ggpubr)
library(ggtext)
library(magick)
library(patchwork)
library(showtext)
library(brms)
library(tidybayes)
library(gt)
library(gnnlab)
library(utils)
library(readxl)
```

```{r}
#| include: false
font_add_google("Noto Sans", family = "ns")
theme_pubr(10, "ns", border = TRUE) |> 
  set_theme()
theme_pubr() |> 
  replace_theme(
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    axis.title.x = element_markdown(family = "ns", margin = margin(0.5, 0, 0.5, 0, "lines")),
    axis.title.y = element_markdown(family = "ns", angle= 90, margin = margin(0, 0.5, 0, 0.5, "lines")),
  )
showtext_auto()
folder = "~/Dropbox/MS_Terada_Greg/"
folder = "./Data/"
fnames = dir(folder, recursive = TRUE)
```


```{r}
ftext = "1800sec"
fname = str_c(folder, str_subset(fnames, ftext)) 
esheets = excel_sheets(fname)
d1800= read_xlsx(fname, range = "A1:BA120") |> 
  pivot_longer(matches("^Y"))
ftext = "900sec"
fname = str_c(folder, str_subset(fnames, ftext)) 
esheets = excel_sheets(fname)
d0900 = read_xlsx(fname, range = "A1:BA120") |> 
  pivot_longer(matches("^Y"))

data = bind_rows(`1800` = d1800, `900` = d0900, .id = "period")

data = data |> 
  mutate(period = as.numeric(period)) |> 
  select(period,
         tau = matches("Sec"),
         name, value) |> 
  group_by(period) |> 
  mutate(N = seq_along(name)) |> 
  ungroup() |> 
  select(-name) |> 
  drop_na()
```

```{r}
xlabel = "Time of exposure (sec)"
ylabel = "Effective quantum yield (&Delta;F/Fm&prime;)"
xbreaks = unique(data$tau)
xbreaks = seq(0, 4200, by = 600)
# 1. Create a label dataframe 
# We take distinct categories and generate the "A: ..." label here
labeldf = data |> expand(period, x = 0, y = 1) |> 
  mutate(
    l = str_glue("({LETTERS[seq_along(period)]}) {period}-sec desiccation ")
  )

rect_df = data |> expand(period, xmin = c(900, 1800), xmax = c(Inf, Inf), ymin = c(-Inf, -Inf), ymax = c(Inf, Inf))

labeldf2 = data |> expand(period, hj = c(0,1,0,1), y = rep(0.15, 4), x = rep(c(0, 3200), 4))

data2 = data |> 
  group_by(period, tau) |> 
  reframe(
    across(value, list(
      m = mean,
      sd = sd,
      med = median,
      mad = mad,
      iqr1 = ~quantile(., prob = 0.25),
      iqr3 = ~quantile(., prob = 0.75)
    ))
  )


p1 = ggplot(data2) + 
  geom_rect(
    aes(
      xmin = xmin, 
      xmax = xmax, 
      ymin = ymin, 
      ymax = ymax),
    data = rect_df, 
    fill = "grey") +
  geom_errorbar(
    aes(x = tau, 
        ymin = value_iqr1,
        ymax = value_iqr3
    ), 
    color = "black",
    linewidth = 0.5) +
  geom_point(aes(x = tau, y = value_med), 
             shape = 21,
             fill = "black",
             stroke = 0) +
  geom_richtext(aes(x = x, y = y, label = l),
                data = labeldf,
                vjust = 1, hjust = 0,
                color = "transparent",
                fill = "transparent",
                text.color = "black",
                family = "ns", 
                size = 3
  ) +
  scale_x_continuous(xlabel, breaks = xbreaks, limits = range(xbreaks)) +
  scale_y_continuous(ylabel, breaks = seq(0, 1, by = 0.25), limits = c(0, 1)) +
  facet_wrap(vars(period), ncol = 1) +
  theme(
    strip.text = element_blank(),
    strip.background = element_blank()
  )

pdfname = "PAM-Kinetics-plot.pdf"
save_plot(pdfname, plot = p1, width = 80, height = 80*1.5)
```



::: {#fig-pam-kinetics}


```{r}
image_read(str_replace(pdfname, "pdf", "png"))
```

The real-time response of the effective quantum yield (&Delta;F/Fm&prime;) in the 
green alga, *Ulva australis* from Nagashima Island, Kagoshima, Japan to 
(A) 1800 seconds of air exposure and 1800 seconds of rehydration and
(B) 900 seconds of air exposure and 1800 seconds of rehydration.
The temperature was 20&deg;C and the humidity was 40%.
Rehydration occurred under the dim light (20 &mu;mol photons m^-2^ s^-1^).
The points are jittered horizontally for clarity (n = 50).
:::

